Avoiding the Global Sort: \break A Faster Contour Tree Algorithm \footnote{The full updated version of this paper is available on the arXiv \cite{rs-mvrms-14}}

### 2016-03-21 paint-contour-tree.tex
done with edits for camera ready version

We think of the dimension $d$, as constant, and assume that $\MM$ is represented in a data structure that allows constant-time access to neighboring simplices
in $\MM$ (e.g.~\cite{BoMa12}). (This is analogous to a doubly connected edge list, but for higher dimensions.)
Observe that $f:\MM \rightarrow \RR$ can be thought of as a $d$-dimensional simplicial complex living in $\RR^{d+1}$, 
where $f(x)$ is the ``height" of a point $x \in \MM$, which is encoded in the representation of $\MM$. 
Specifically, rather than writing our input as $(\MM,f)$, we abuse notation and typically just write $\MM$ to denote the lifted complex.

@@

@@

\begin{lemma} \label{lem:full} Suppose vertex $v$ is connected to a component $\PP$ of $\MM^+_v$
by an edge $e$ which is currently colored $c$. Either all edges in $\PP$ are currently colored $c$, or there
exists a critical vertex $w \in \PP$ fully touched by $c$ and touched by another color.
\end{lemma}

>>>>

We think of the dimension $d$, as constant, and assume that $\MM$ is represented in a data structure that allows constant-time access to neighboring simplices
in $\MM$ (e.g. incidence graphs~\cite{Ed87}). (This is analogous to a doubly connected edge list, but for higher dimensions.)
Observe that $f:\MM \rightarrow \RR$ can be thought of as a $d$-dimensional simplicial complex living in $\RR^{d+1}$, 
where $f(x)$ is the ``height" of a point $x \in \MM$, which is encoded in the representation of $\MM$. 
Specifically, rather than writing our input as $(\MM,f)$, we abuse notation and typically just write $\MM$ to denote the lifted complex.

@@

\medskip
\noindent
{\bf Color list $\col(v)$:} For each point $v$, we maintain $\col(v)$ as a simple list.
In addition, we will maintain another (sub)list of colors $L$ such that $\forall c \in L$,
$v$ is \emph{not} the highest vertex in $T(\rep(c))$. Note that given $c \in \col(v)$
and $\rep(c)$, this property can be checked in constant time. If $c$ is ever removed
from this sublist, it can never enter it again.

For notational convenience, we will not explicitly
maintain this sublist. We simply assume that, in constant
time, one can determine (if it exists) an arbitrary color $c \in \col(v)$ such that $v$ is not the highest vertex in $T(\rep(c))$.

@@

\begin{lemma} \label{lem:full} Suppose vertex $v$ is connected to a component $\PP$ of $\MM^+_v$
by an edge $e$ which is currently colored $c$. At all times: either all edges in $\PP$ are currently colored $c$, or there
exists a critical vertex $w \in \PP$ fully touched by $c$ and touched by another color.
\end{lemma}
